The origin of cosmic-ray electrons in cluster outskirts

### 2012-01-27 paper.tex
text update

\subsection{CR protons}
The CR proton distribution function is given by
\begin{equation}
  \label{eq:f_p}
  f_\p(p_\p) = \frac{\dd^2 N_\p}{\dd p_\p\,\dd V} = 
  C\, p_\p^{-\alpha}\,\theta(p_\p-q)\,, 
\end{equation}
and $p_\p = P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$.  Here $q$ is the normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization in units of density.

\del{From the galaxy cluster simulations
we determine the adiabatic invariant Lagrangian amplitude of the
spectrum $\tilde{C}_0$, where the adiabatic changes of the gas density
$\rho$ change the normalization according to
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
and the convenient unitless redefinition of the amplitude is given by 
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}}

From our simulated galaxy clusters we derive for each snapshot at time
$t$ the injected CR proton distribution function $f_\p$. We select
particles in the outskirts of clusters where the gas densities are low
and hence the Coulomb cooling of the CR protons small. The CR proton
distribution function from the simulations is parameterized in terms of
adiabatic invariant momentum cutoff $q_0$ and the adiabatic invariant
Lagrangian amplitude of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}
The injected distribution function is calculated as the change in
normalization between each snapshot, $f_\rmn{inj,p}(p_\p) = \Delta
C\,p_\p^{-\alpha}$, and is derived through the formalism developed
in \cite{2007A&A...473...41E} and \cite{2008A&A...481...33J}:
\begin{eqnarray}
  \label{eq:deltaC}
  \Delta C(t+\Delta t) &=& C(t) 
  \frac{\Delta \eps_\CR(t)-T_\p(q(t))\,\Delta n_\CR(t)}
       {\eps_\CR(t) - T_\p(q(t))\,n_\CR(t)}\,,\quad\rmn{where}\nonumber\\
       \Delta X(t) &=&  X(t+\Delta t) -  X(t)\,. 
\end{eqnarray}
We fix the time between each snapshot $\Delta t$ to $100$~Myrs, which
is smaller then the loss timescale of CR electrons in cluster
outskirts. The CR proton number density is given by
\begin{equation}
  \label{eq:ncr}
  n_{\CR} = \int_0^\infty \dd p_\p\, f_\p(p_\p) =
  \frac{C\, q^{1-\alpha}}{\alpha-1}\,,
\end{equation}
provided $\alpha >1$. The kinetic energy density of the CR
proton population is
\begin{eqnarray}
  \label{eq:epscr}
  \eps_\CR &=& \int_0^\infty  \dd p_\p\, f_\p(p_\p) \,T_{\rmn
    p}(p_\p)=\frac{C\, m_\p\,c^2}{\alpha-1} \, \times
  \nonumber \\
  && \left[\frac{1}{2}
    \, \B_{\frac{1}{1+q^2}} \left(
    \frac{\alpha-2}{2},\frac{3-\alpha}{2}\right) + q^{1-\alpha}
    \left(\sqrt{1+q^2}-1 \right) \right] \,,
\end{eqnarray}
where $T_\p(p_\p) = (\sqrt{1+p_\p^2} -1)\, m_\p\,c^2$ is the kinetic energy
of a CR proton. $\B_x(a,b)$ denotes the incomplete Beta-function, and
$\alpha>2$ is assumed.

>>>>

\begin{table}
\caption{Cluster sample.}
\begin{tabular}{l l l l r r r}
\hline
\hline
Cluster & sim.'s & dyn. state$^{(1)}$ & $M_{\rm vir}^{(2)}$ & $\rvir^{(2)}$ & $kT_{\rm vir}^{(3)}$ & $\Delta t$$^{(1=4)}$\\
& & & [$\rmn{M}_\odot$] & [Mpc] & [keV] & \\
\hline
1  & g8a  & CC    & $2.6\times 10^{15}$ &   2.9~~ & 13.1 \\
2  & g1a  & CC    & $1.9\times 10^{15}$ &   2.5~~ & 10.6 \\
3  & g72a & PostM & $1.6\times 10^{15}$ &   2.4~~ & 9.4  \\
4  & g51  & CC    & $1.5\times 10^{15}$ &   2.4~~ & 9.4  \\

5  & g1b  & M     & $5.2\times 10^{14}$ &   1.7~~ & 4.7  \\
6  & g72b & M     & $2.2\times 10^{14}$ &   1.2~~ & 2.4  \\
7  & g1c  & M     & $2.0\times 10^{14}$ &   1.2~~ & 2.3  \\
8  & g8b  & M     & $1.5\times 10^{14}$ &   1.1~~ & 1.9  \\
9  & g1d  & M     & $1.3\times 10^{14}$ &   1.0~~ & 1.7  \\

10 & g676 & CC    & $1.3\times 10^{14}$ &   1.0~~ & 1.7  \\
11 & g914 & CC    & $1.2\times 10^{14}$ &   1.0~~ & 1.6  \\
12 & g1e  & M     & $9.1\times 10^{13}$ &  0.93   & 1.3  \\
13 & g8c  & M     & $8.5\times 10^{13}$ &  0.91   & 1.3  \\
14 & g8d  & PreM  & $7.8\times 10^{13}$ &  0.88   & 1.2  \\
\hline
\end{tabular}  \begin{quote}
 Notes:\\ (1) The dynamical state has been classified through a combined
 criterion invoking a merger tree study and the visual inspection of the X-ray
 brightness maps. The labels for the clusters are M--merger, PostM--post
 merger (slightly elongated X-ray contours, weak cool core (CC) region developing),
 PreM--pre-merger (sub-cluster already within the virial radius), CC--cool
 core cluster with extended cooling region (smooth X-ray profile).  (2) The
 virial mass and radius are related by $M_\Delta(z) = \frac{4}{3} \pi\,
 \Delta\, \rho_\rmn{crit}(z) R_\Delta^3 $, where $\Delta=200$ denotes a
 multiple of the critical overdensity $\rho_\rmn{crit}(z) = 3 H (z)^2/ (8\pi
 G)$.  (3) The virial temperature is defined by $kT_{\rm vir} = G M_{\rm vir} \,
 \mu\, m_\p / (2 R_{\rm vir})$, where $\mu$ denotes the mean molecular weight.
\end{quote}
\label{tab:cluster_sample}
\end{table}

INSERT CLUSTER TABLE WITH DELTA T HERE, USE 100 MYR, AND THE LOG A

\subsection{CR protons}
For each SPH fluid element we attach a distribution function of CR protons. It is represented by a one-dimensional isotropic distribution function \footnote{The three dimensional distibution function is given by $4\pi \int f(p) p^ 2 \dd p$.} given by
\begin{equation}
  \label{eq:f_p}
  f_\p(p_\p) = \frac{\dd^2 N_\p}{\dd p_\p\,\dd V} = 
  C\, p_\p^{-\alpha}\,\theta(p_\p-q)\,, 
\end{equation}
and $p_\p = P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$.  Here $q$ is the mass normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization of the distribution function in units of density.

\del{From our simulated galaxy clusters we derive for each snapshot at time
$t$ the injected CR proton distribution function $f_\p$. We select
particles in the outskirts of clusters where the gas densities are low
and hence the Coulomb cooling of the CR protons small. }

The CR proton
distribution function from the simulations is convinently parameterized in terms of
adiabatic invariant momentum cutoff $q_0$ and the adiabatic invariant
Lagrangian amplitude of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}

From our simulated galaxy clusters we derive for each snapshot at time
$t$ a CR proton distribution function $f_\p$.

\del{We select
particles in the outskirts of clusters where the gas densities are low
and hence the Coulomb cooling of the CR protons small.}

The injected distribution function is calculated as the change in
CR normalization between a snapshot at time $t$ and an earlier time 
$t-\Delta t$:
\begin{eqnarray}
  \label{eq:finj}
f_\rmn{inj,p}(t,p_\p) &=& \Delta C(t)\,\frac{\rho}{m_\p}\,\rmn{AD(\rho)} \,p_\p^{-\alpha}\qquad\rmn{where}\\
  \Delta  C(t) &=& C(t) - C(t-\Delta t)\,
\end{eqnarray}
and the adiabatic cooling term $\rmn{AD}(\rho)\equiv\left(\rho/\rho_0\right)^{-1/3}$.

We fix the time between each snapshot $\Delta t$ to $100$~Myrs, which
is smaller then the loss timescale of CR electrons in cluster
outskirts.

\del{The CR proton number density is given by
\begin{equation}
  \label{eq:ncr}
  n_{\CR} = \int_0^\infty \dd p_\p\, f_\p(p_\p) =
  \frac{C\, q^{1-\alpha}}{\alpha-1}\,,
\end{equation}
provided $\alpha >1$. The kinetic energy density of the CR
proton population is
\begin{eqnarray}
  \label{eq:epscr}
  \eps_\CR &=& \int_0^\infty  \dd p_\p\, f_\p(p_\p) \,T_{\rmn
    p}(p_\p)=\frac{C\, m_\p\,c^2}{\alpha-1} \, \times
  \nonumber \\
  && \left[\frac{1}{2}
    \, \B_{\frac{1}{1+q^2}} \left(
    \frac{\alpha-2}{2},\frac{3-\alpha}{2}\right) + q^{1-\alpha}
    \left(\sqrt{1+q^2}-1 \right) \right] \,,
\end{eqnarray}
where $T_\p(p_\p) = (\sqrt{1+p_\p^2} -1)\, m_\p\,c^2$ is the kinetic energy
of a CR proton. $\B_x(a,b)$ denotes the incomplete Beta-function, and
$\alpha>2$ is assumed.}
